A survey of HTTP caching options on the open Web

### 2015-01-14 sanity/paper1.tex
Smaller proofreading
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@980 43ef1190-90fd-4dd7-a769-d0121f02b80e

\newcommand{\rdfterm}[1]{\texttt{#1}}
\newcommand{\pvalue}{\textit{p}-value}
\newcommand{\httph}[1]{\texttt{#1}}
\newcommand{\todo}[1]{\ensuremath{^{\textrm{\tiny{TODO}\normalsize}}}\footnote{\textbf{TODO:}~#1}}

@@

For each unique \httph{Server} header, we found the ones where
\emph{all} responses had a freshness time or a promising feature. For
the former, this amounted to 22 servers, tabulated
in~\ref{tab:servers}. 70 servers always responded with promising
features. As we can see, apart from well-known Virtuoso and
Callimachus servers, as well as the Perl modules RDF::LinkedData and
RDF::Endpoint, which is partly developed by and runs on a server
operated by this author, most headers tells very little about the
RDF-specific parts of the underlying server implementation. A quick
inspection of all \httph{Server} headers confirmed that few reveal any
further detail.

% latex table generated in R 2.15.1 by xtable 1.5-6 package
% Mon Jan  5 11:27:35 2015
\begin{table}[ht]
  \caption{\httph{Server} headers for hosts that enabled a freshness
    lifetime to be computed for all requests.}\label{tab:servers}
\begin{center}
\begin{tabular}{rl}
  \hline
 & x \\ 
  \hline
1 & DFE/largefile \\ 
  2 & git\_frontend \\ 
  3 & nginx/1.3.9 \\ 
  4 & thin 1.6.0 codename Greek Yogurt \\ 
  5 & Oracle-Application-Server-10g/10.1.3.4.0 Oracle-HTTP-Server OracleAS-Web-Cache-10g/10.1.2.3.1 (N;ecid=72058133508068653,0) \\ 
  6 & Oracle-Application-Server-10g/10.1.3.4.0 Oracle-HTTP-Server OracleAS-Web-Cache-10g/10.1.2.3.1 (N;ecid=72057948824487974,0) \\ 
  7 & TwistedWeb/8.2.0 \\ 
  8 & RDF::Endpoint/0.07 \\ 
  9 & Jetty(6.1.26) \\ 
  10 & nginx/1.6.1 \\ 
  11 & Jigsaw/2.3.0-beta3 \\ 
  12 & Apache/2.2.9 (Win32) PHP/5.2.6 \\ 
  13 & Apache/2.4.10 (Unix) mod\_fcgid/2.3.9 \\ 
  14 &  \\ 
  15 & GFE/2.0 \\ 
  16 & RDF::LinkedData/0.70 \\ 
  17 & Apache/2.2.17 (Unix) mod\_wsgi/3.3 Python/2.6.6 \\ 
  18 & Virtuoso/07.10.3211 (Linux) i686-generic-linux-glibc212-64  VDB \\ 
  19 & Apache/2.2.24 (Unix) mod\_ssl/2.2.24 OpenSSL/0.9.8y \\ 
  20 & Apache/2.2.22 (Fedora) \\ 
  21 & INSEE \\ 
  22 & GitHub.com \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

@@

In both the cases of standard-compliant freshness time and for the
other promising features, the test reports \pvalue = 0.0001. We can
conclude that it is highly likely that some servers are better at
exposing cache headers than others. Unfortunately, since most
\httph{Server} headers only report generic values, little can be
learnt about these implementations. We note, however, that DBPedia
exposes standards-compliant freshness time of 604800 seconds (i.e. 1
week) for both LOD and SPARQL endpoints.

@@

\subsection{Distribution of freshness time}

We obtained a successful response from a total 2965 information
resources, either with SPARQL results or RDF data. A successful
response is rather tightly defined, not only must there be a
successful HTTP response after redirects are resolved, the response must also
return a valid RDF media type (unless it is a SPARQL result) and the
RDF must parse into an RDF model.

@@

We next consider the simple heuristic freshness lifetime. This is
implemented as suggested in RFC7234, section~4.2.2. This is based on a
fraction of the time lapsed between the current time and the
modification time given in the \httph{Last-Modified} header. This
approach therefore still requires the Web server to be cooperative to
be successful, but commonly, Web servers can track this, based e.g. on
the file modification time of the underlying file system.

@@

We were able to compute a heuristic lifetime for 554 resources, a
larger number than standards-compliant resources. In
Figure~\ref{fig:heuristicall}, we see that the distribution of
lifetimes are radically different from the case in
Figure~\ref{fig:hardall}. In this case, we may cache many resources
for months. Only a handful of resources had changed in the last
minutes. Since this is based on actual times since last modifications,
this suggests that many resources should have had explicit cache
headers with very long lifetimes. This is supported by
DyLDO\cite{dyldo2}, which concludes that:
\begin{quote}
[...] We found that 62.2\% of documents didn’t change over the six
months and found that 51.9\% of domains were considered static.
\end{quote}
This agrees well with that 60\% of the simple heuristic lifetimes are
in the month range.

@@

Based on the resulting headers, we first let libwww calculate the
freshness lifetime in seconds if it was possible, and recorded
that. Then, we also let it calculate a \emph{heuristic} freshness
time based on a suggestion in RFC7234, section~4.2.2, and recorded that.

>>>>

\newcommand{\rdfterm}[1]{\texttt{#1}}
\newcommand{\pvalue}{\textit{p}-value\ }
\newcommand{\httph}[1]{\texttt{#1}}
\newcommand{\todo}[1]{\ensuremath{^{\textrm{\tiny{TODO}\normalsize}}}\footnote{\textbf{TODO:}~#1}}

@@

For each unique \httph{Server} header, we found the ones where
\emph{all} responses had a freshness time or a promising feature. For
the former, this amounted to 22 servers, see 
Table~\ref{tab:servers}. 70 servers always responded with promising
features. As we can see, apart from well-known Virtuoso and
Callimachus servers, as well as the Perl modules RDF::LinkedData and
RDF::Endpoint, which is partly developed by and runs on a server
operated by this author, most headers tells very little about the
RDF-specific parts of the underlying server implementation. A quick
inspection of all \httph{Server} headers confirmed that few reveal any
further detail.

% latex table generated in R 2.15.1 by xtable 1.5-6 package
% Mon Jan  5 11:27:35 2015
\begin{table}[ht]
  \caption{\httph{Server} headers for hosts that enabled a freshness
    lifetime to be computed for all requests.}\label{tab:servers}
\begin{center}
\begin{tabular}{rl}
  \hline
1 & DFE/largefile \\ 
  2 & git\_frontend \\ 
  3 & nginx/1.3.9 \\ 
  4 & thin 1.6.0 codename Greek Yogurt \\ 
  5 & Oracle-Application-Server-10g/10.1.3.4.0 Oracle-HTTP-Server [\ldots]\\
  6 & Oracle-Application-Server-10g/10.1.3.4.0 Oracle-HTTP-Server [\ldots]\\
  7 & TwistedWeb/8.2.0 \\ 
  8 & RDF::Endpoint/0.07 \\ 
  9 & Jetty(6.1.26) \\ 
  10 & nginx/1.6.1 \\ 
  11 & Jigsaw/2.3.0-beta3 \\ 
  12 & Apache/2.2.9 (Win32) PHP/5.2.6 \\ 
  13 & Apache/2.4.10 (Unix) mod\_fcgid/2.3.9 \\ 
  14 &  \\ 
  15 & GFE/2.0 \\ 
  16 & RDF::LinkedData/0.70 \\ 
  17 & Apache/2.2.17 (Unix) mod\_wsgi/3.3 Python/2.6.6 \\ 
  18 & Virtuoso/07.10.3211 (Linux) i686-generic-linux-glibc212-64  VDB \\ 
  19 & Apache/2.2.24 (Unix) mod\_ssl/2.2.24 OpenSSL/0.9.8y \\ 
  20 & Apache/2.2.22 (Fedora) \\ 
  21 & INSEE \\ 
  22 & GitHub.com \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

@@

In both the cases of standard-compliant freshness time and for the
other promising features, the test reports \pvalue = 0.0001. We can
conclude that it is highly likely that some servers are better at
exposing cache headers than others. Unfortunately, since most
\httph{Server} headers only contain generic values, little can be
learnt about these implementations. We note, however, that DBPedia
exposes standards-compliant freshness time of 604800 seconds (i.e. 1
week) for both LOD and SPARQL endpoints.\todo{hall of fame?}

@@

\subsection{Distribution of freshness lifetime}

We obtained a successful response from a total 2965 information
resources, either with SPARQL results or RDF data. A successful
response is rather strictly defined, not only must there be a
successful HTTP response after redirects are resolved, the response
must also return a valid RDF media type (unless it is a SPARQL result)
and the response must parse into an RDF model.

@@

We next consider the simple heuristic freshness lifetime. This is
implemented as suggested in RFC7234, Section~4.2.2. This is based on a
fraction of the time lapsed between the current time and the
modification time given in the \httph{Last-Modified} header. This
approach therefore still requires the Web server to be cooperative to
be successful, but commonly, Web servers can track this, based e.g. on
the file modification time of the underlying file system.

@@

We were able to compute a heuristic lifetime for 554 resources, a
larger number than standards-compliant resources. In
Figure~\ref{fig:heuristicall}, we see that the distribution of
lifetimes are radically different from the case in
Figure~\ref{fig:hardall}. In this case, we may cache many resources
for months. Only a handful of resources had changed in the last
minutes. Since this is based on actual times since last modifications,
this suggests that many resources should have had explicit cache
headers with very long lifetimes. This is supported by
DyLDO\cite{dyldo2}, which concludes that:
\begin{quote}
[\ldots] We found that 62.2\% of documents didn’t change over the six
months and found that 51.9\% of domains were considered static.
\end{quote}
This agrees well with that 60\% of the simple heuristic lifetimes are
in the month range.

@@

Based on the resulting headers, we first let libwww calculate the
freshness lifetime in seconds if it was possible, and recorded
that. Then, we also let it calculate a \emph{heuristic} freshness
time based on a suggestion in RFC7234, Section~4.2.2, and recorded that.
